---
title: "MBIO630_GroupProject_MeasuringReefs"
author: "Kyleigh Kuball"
date: "`r format(Sys.Date())`" 
output: 
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sciplot)
library(vegan)
library(plotrix)
library(plyr)
```

## Introduction 

Satellite imagery is a powerful tool that has been used to study animal populations, tree cover, and is slowly expanding to use in other areas of study (LaRue, Stapleton, & Anderson 2017). Satellite imagery has been used to study fish populations as well as coral reef structure (Madin & Foley 2021; Madin & Madin 2015). Further studies will continue to use satellite imagery and spatial resolution to measure features pertinent to areas of research (Madin & Madin 2015). Moving forward it is important to understand the how satellite imagery spatial resolution affects measurements of research data within research sites. This study looked at three methods of measurement (GPS, transect, and satellite data) in regards to the benefits and disadvantages of each method. 

## Materials & Methods

This study was conducted in Kāneʻohe Bay, Hawaiʻi at reef patches 20 and 22. Transect and gps data was taken by two groups of 3 snorkelers swimming the perimeter of each patch reef. Transects were laid out end over end to measure the perimeter while one person in the group of 3 followed with a GPS. Satellite imagery was taken from multiple dates via Google Earth Pro (v.7.3.4.8642) and Planet Explorer. Perimeter within the satellite data was taken using the polygon tool in both instances and then calculating the area of a circle (((perimeter/(2*PI()))^2)*PI()) via Excel(v. 16.65). All data was then compiled into R Studio (v. 2022.07.1) for analysis. 

## Results

Overall, Figure 1 illustrates that GPS has a higher estimate of area, imagery has a lot of outliers and is quite variable, and the transect data appears to have the least outliers and smallest average. However, the average between imagery and transect appear to be similar. Figure 2 illustrates the significant correlation (t = 9.43, p = <2e-16) that the better the imagery resolution, the smaller the estimate for the patch reef area. It appears that the area is estimated to be greater when the resolution is poor. In relation, a significant correlation was found between the resolution and the perimeter (t = 8.493, p = 2.32e-15); Figure 3 illustrates this correlation that the greater the resolution, the greater the estimate of the perimeter. Figure 4 shows that the reef is increasing in area as time increases. The linear model run shows that this increase shown in Figure 4 is statistically significant (t = 3.209, p = 0.00151).

Imagery resolution (t = -2.780, p = 0.00588) and year (t = -2.230, p = 0.02673) both had a significant correlation with the area and there was a significant correlation between imagery resolution and year (t = 2.785, p = 0.00579). An Analysis of Variance revealed that year may not be a significant variable when isolated from imagery resolution however when the drop1 function was used, no parameters were suggested to be dropped. 

The relationship between reef number and area is significant (t = -2.95, p =0.0059) and there is a negative correlation between the two due to reef 22 having a decline in area when compared to reef 20. There appears to be no significant interaction between imagery resolution and reef number, but there is a significant correlation between imagery resolution and area (t = 3.96, p = 0.00039). This means that the variation in the imagery resolution is not significantly correlated to the reef number but is significant when calculating the area at either reef. 

```{r data import, echo=TRUE, results='hide', include=FALSE}

data<-read.csv("data/MeasuringReefs_Master (1).csv", header = TRUE)

names(data)
head(data)
data$year<-substr(data$date,0,4) # add a column for year only

data
```


```{r data explore-1, echo=F, results='hide', fig.height=5, fig.width=5, fig.align='center', fig.cap="\\label{fig:figs}plotting example" }

par(mar = c(9,2,5,2))
boxplot(area_m2~method, data=data, na.ignore=TRUE, col="lightgrey",
     ylab="Patch reef area (m^2)", 
     xlab="Method",
     main="Figure 1. Area versus method")
mtext(side=3, paste(
  strwrap(
    'The different methods of measuring the area of the patch reefs  20 and 22 is shown on the x-axis and the patch reef area is shown on the y-axis',
    width = 50), collapse = "\n"), line =-20)
```


```{r, echo = F}
## on area
par(mar = c(9,3,5,2))
plot(area_m2~imagery_resolution_m, data=data, col="slategrey",
     xlim=c(0,10),
     xlab="Imagery spatial resolution (m)", 
     ylab="Patch reef area (m^2)",
     main="Figure 2. Area versus imagery resolution")
mtext(side=3, paste(
  strwrap(
    'The imagery spatial resolution in meters is shown on the x-axis and the area of patch reef estimated (m^2) is shown on the y-axis. The patch reef area increases with the spatial resolution.',
    width = 50), collapse = "\n"), line =-20)     

res.area.lm=lm(area_m2 ~ imagery_resolution_m, data=data)
abline(res.area.lm, col = "black")
summary(res.area.lm)
```

```{r, echo = F}
## on perimeter
par(mar = c(9,3,5,2))
plot(perimeter_m~imagery_resolution_m, data=data, col="slategrey",
     xlim=c(0,10),
     xlab="Imagery spatial resolution (m)", 
     ylab="Patch reef perimeter (m)", 
     main="Figure 3. Perimeter versus imagery resolution")
mtext(side=3, paste(
  strwrap(
    'The spatial resolution is on the x-axis and the patch reef perimeter on the y-axis. Similar to Figure 2, as the spatial resolution increases so does the estimate of the patch reef area.',
    width = 50), collapse = "\n"), line =-20) 

res.perim.lm=lm(perimeter_m ~ imagery_resolution_m, data=data)
abline(res.perim.lm, col = "black")
summary(res.perim.lm)
```

```{r, echo = F}
# Effect of date (i.e., do patch reefs measurably grow or shrink over time?)
par(mar = c(9,3,5,2))
plot(area_m2~as.numeric(year), data=data, col="slategrey",
     xlab="Year", 
     ylab="Patch reef area (m^2)", 
     main="Figure 4. Area versus year")
mtext(side=3, paste(
  strwrap(
    'The year is on the x-axis and the patch reef area is shown on the y-axis. As the time in years increases, so does the patch reef area.',
    width = 50), collapse = "\n"), line =-20) 

date.area.lm=lm(area_m2 ~ as.numeric(year), data=data)
abline(date.area.lm, col = "black")
summary(date.area.lm)

```



Further analysis:

```{r data analysis-1, echo=F}

data$year <- as.numeric(data$year) 
res.year.lm=lm(area_m2 ~ imagery_resolution_m * year, data=data) # run interaction model
summary(res.year.lm)      # overall model is significant
anova(res.year.lm)        # run ANOVA to look at predictor variables in isolation
drop1(res.year.lm, test="F")  # use drop1 function to see which parameter can be dropped (test="F" adds type 
                          # II ANOVA

```

Analysis on reef identity (reef number):

```{r data analysis-2, echo=F}

res.reefID.lm=lm(area_m2 ~ imagery_resolution_m * reef_no,  data= na.omit(data) )
summary(res.reefID.lm)  
              # The relationship between reef number and area is significant (t = -2.95, p =0.0059). There is a negative correlation between the two meaning that reef 22 likely has a decline in area when compared to reef 20.  There appears to be no significant interaction between imagery resolution and reef number, but there is a significant correlation between imagery resolution and area (t = 3.96, p = 0.00039). This means that the variation in the imagery resolution is not significantly correlated to the reef number but is significant when calculating the area at either reef. 

```

    
## Discussion

In this study, we looked at how spatial resolution affects subject areas of research. Overall, we found that resolution can greatly affect the perimeter and area of a site, with imagery data being far more variable among individuals while the field/transect data seemed to have significantly less outliers than the imagery data. The measurements of each reef tended to increase as the resolution become poorer, this was likely due to a difficulty in determining the edges of each patch reef. An interesting find in this data set was that given that the spatial imagery resolution gets smaller and better with time, the size of the patch reefs would be assumed to decrease as time increases based on the figures in this study. Because of the resolution getting better, if the reef was not growing or changing, the edges would be easier to identify so the area would decrease (Figure 2 & 3). Considering that it is increasing, there is likely substantial growth since the resolution increasing should be working against the measurements taken. 

In all, there is a balance between imagery measurements and field measurements. Imagery measurements are far less time consuming and costly than field data, and more dates and times may be available to compare to. However, the field/transect data tends to be a lot more accurate with far less outliers than the imagery data which would be important in certain research projects. The imagery platform trade-off seems to be the difference in price with the difference in resolution. The lower resolution platforms definitely had more variability among groups and was far less accurate, however that is the least expensive method to monitor the reef long term. As the resolution goes up, the price increases and it starts to be a balance of if it is less expensive to be in the field or use the imagery data. There is more data available in low resolution than there is in high resolution. High resolution tends to be expensive and sometimes the clouds are covering the area of interest. With low resolution, pictures are taken more often and so cloud cover isn't as big of a deal and it is less expensive. However, the area estimate variability goes up by a lot as the resolution becomes poorer. 

## References

LaRue, M. A., Stapleton, S., & Anderson, M. (2017). Feasibility of using high‐resolution satellite imagery to assess vertebrate wildlife populations. Conservation biology, 31(1), 213-220.

Madin, E. M., & Foley, C. M. (2021). The Shift to a Bird's-Eye View: Remote sensing technologies allow researchers to track small changes on a large scale and enable studies of far-flung places from the comfort and safety of home. American Scientist, 109(5), 288-296.

Madin, J. S., & Madin, E. M. (2015). The full extent of the global coral reef crisis. Conservation Biology, 29(6), 1724-1726.
